### 2021-02-02 written/eLifeRevision/Bittner_eLife2020.tex
latest tex

\includegraphics[scale=0.32]{figures/fig1/fig1.pdf}

@@

\begin{figure}
\vspace{-1cm}
\begin{center}
\includegraphics[scale=0.2]{figures/fig2/fig2.pdf}
\end{center}
\vspace{-1cm}
\caption{\small Emergent property inference of a stochastic stabilized supralinear network.
\textbf{A}.  Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).   
Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_b$.
Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
\textbf{B}. Responses of the deterministic model ($\bm{\epsilon}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
\textbf{C}. Paradoxical response of the stochastic model to a small increase in input to the P-population.  
\textbf{D}. EPI posterior of noise parameters $\mathbf{z}$ conditioned on realistic E-population Fano factors.
The posterior predictive distribution is show on the bottom-left.
and the mode of the distribution is starred.
\textbf{E}. (Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of the posterior.
Each gray dot is a choice of $\sigma_P$, for which a constrained mode $z^*(\sigma_P, P)$ is chosen.
The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
(Bottom) Such sensitive dimensions of the Hessian (dots) are significantly more sensitive than randomly chosen dimensions (box and whiskers).
\textbf{F}. The Fano factor of the E-population is strongly correlated with each other neuron-type population.
\textbf{G}. Mean and standard deviation (across EPI posterior) of Fano factor of each neuron-type population at each level of contrast.
}
 \label{fig:V1}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.48]{figures/fig3/fig3.pdf}
\end{center}
\caption{\small 
\textbf{A}. Rapid task switching behavioral paradigm (see text). 
\textbf{B}. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
\textbf{C}. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice of simulations (D).
\textbf{D}. Simulations of an SC network from the EPI distribution with 75\% accuracy in each task.  Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (optogenetic strength $\gamma=0.7$).  Shading indicates standard deviation across trials.
\textbf{E}. Difference in performance of each task during inactivation.  Inactivation level $\gamma$ scales from no inactivation (0) to full inactivation (1).  We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
\textbf{F}. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
}
\label{fig:SC}
\end{figure}

@@

Here, we are interested in the scalability of EPI in number of parameters ($|\mathbf{z}|$).  We consider rank-2 RNN with N neurons of connectivity
\begin{equation}
 W = UV^\top + g\chi
 \end{equation}
 and dynamics
 \begin{equation}
 \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x}
 \end{equation}
where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix}$, $\mathbf{u}_1, \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $g=0.01$.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=1.4]{figures/fig4/fig4.pdf}
\end{center}
\caption{\small 
\textbf{A}. Recurrent neural network.
\textbf{B}. EPI scales with $z$ to high dimensions. 
Convergence definitions: 
EPI (blue) - satisfies all moment constraints, 
SNPE (orange)- produces at least 2/$n_{\text{train}}$ parameter samples are in the bounds of emergent property (mean +- 0.5), 
and SMC-ABC (red) - 100 particles with $\epsilon < 0.5$ are produced.
\textbf{C}. Posterior predictive distributions of EPI (blue), SNPE (orange), and SMC-ABC (red). 
Gray star indicates emergent property mean, and gray dashed lines indicate two standard deviations corresponding to the variance constraint.
For $N <= 6$ where SMC-ABC converges, samples are not diverse (path degeneracies).  
For $N >= 25$, SNPE does not produce a posterior approximation yielding parameters with simulations near $x_0$. 
\textbf{D}. Simulations of network parameters resulting from each method ($\tau=100ms$).  Each trace corresponds to simulation of one $z$.  
\textbf{E}. Ratio of obtained samples producing stable amplification.
}
\label{fig:LRRNN}
\end{figure}

@@

However, this computation has cubic complexity in dimensionality for fully connected layers.  
By restricting our layers to normalizing flows \cite{rezende2015variational, papamakarios2019normalizing} -- bijective functions with fast log determinant Jacobian computations, which confer a fast calculation of the sample log probability.
Fast log probability calculation confers efficient optimization of the maximum entropy objective (see Section \ref{methods_AL_opt}).
%Most of our analyses use either a planar flow \cite{rezende2015variational} or real NVP \cite{dinh2017density}, which have proven effective in our architecture searches.  
%Planar flow architectures are specified by the number of planar bijection layers used, while real NVP architectures are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
We use the Real NVP \cite{dinh2017density} normalizing flow class, because its coupling architecture confers both fast sampling (forward) and fast log probability evaluation (backward).
Fast probability evaluation in turn facilitates fast gradient and Hessian evaluation of log probability throughout parameter space.
Glow permutations were used in between coupling stages \cite{kingma2018glow}.
This is in contrast to autoregressive architectures \cite{papamakarios2017masked, kingma2016improved}, in which only forward or backward passes are efficient.
Normalizing flow architectures for deep probability distributions used in EPI are specified by the number of coupling stages, neural network layers and units per layer for conditioning functions.

\subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
\begin{equation} \label{eq:AL}
L(\bm{\theta}; \bm{\eta}_{\text{opt}}, c) = -H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top R(\bm{\theta}) + \frac{c}{2}||R(\bm{\theta})||^2
\end{equation}
where $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient. 
These Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
Deep neural network weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.

@@

The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations. 
A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
The p-value of $\mathbb{E}[||R(\bm{\theta}_{k+1})||] > \gamma \mathbb{E} \left[||R(\bm{\theta}_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu}_{\text{opt}})$ given a batch size $n$.
Throughout the study, $\beta = 4.0$, $\gamma = 0.25$, and the batch size was a hyperparameter, which varied according to the application of EPI.

@@

To assess whether the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ produces the emergent property, we defined a hypothesis testing convergence criteria.  
The algorithm has converged when a null hypothesis test of constraint violations $R(\bm{\theta})_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.  
This significance threshold is adjusted through Bonferroni correction according to the number of constraints $m$.  
The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made from $k$ resamplings of $\mathbf{z}$ from a finite sample of size $n$ taken at the end of the augmented Lagrangian epoch.

@@

\subsubsection{Example: 2D LDS}\label{methods_2DLDS}
To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
\begin{equation} 
\tau \frac{d\mathbf{x}}{dt} = A\mathbf{x}
\end{equation}
with
\begin{equation}
A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
\end{equation}
To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(\mathbf{x})$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system. 
 $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
\begin{equation}
 \mathbb{E}\left[T(\mathbf{x}) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}.
 \end{equation} 
\begin{figure}
\begin{center}
\includegraphics[scale=0.58]{figures/figLDS1/figLDS1.pdf}
\end{center}
\begin{flushleft}
\caption{\small (LDS1): A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
B. The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.  Dashed lines indicate the parameter axes. 
C. Entropy throughout the optimization.  
At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.  
D. Emergent property moments throughout optimization.  
At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
\end{flushleft}
\label{fig:LDS1}
\end{figure}

@@

For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three coupling layers of affine transformations parameterized by two-layer neural networks of 25 units per layer.
The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$. 
We did not include $g_{\text{synA}} < 0.01$, since conductances that low make the circuit simulations numerically unstable.
We used an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, set $\nu = 0.25$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce a gaussian approximation to samples returned from an initial ABC search.
This initialization had much greater entropy and a different emergent property than the returned EPI posterior.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.85]{figures/figSC1/figSC1.pdf}
\end{center}
\caption{\small (SC1): Connectivity parameters of EPI distributions versus task accuracies.  $\beta$ is slope coefficient of linear regression, $r$ is correlation, and $p$ is the two-tailed p value.
}
\label{fig:SC1}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=1.3]{figures/figSC2/figSC2.pdf}
\end{center}
\caption{\small (SC2):  
A. Invariant eigenvectors of connectivity matrix $W$.
B. Eigenvalues of connectivities of EPI distribution versus task accuracies.}
\label{fig:SC2}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{figures/figSC3/figSC3.pdf}
\end{center}
\caption{\small (SC3): 
A. Connectivity eigenvalues of EPI parameter distribution colored by Pro task accuracy. 
B. Same for Anti task.
}
\label{fig:SC3}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.75]{figures/figSC4/figSC4.pdf}
\end{center}
\caption{\small (SC4):  Scatters of the effect of delay period inactivation in each task with task accuracy.  Plots are shown at an opto strength of 0.8.
}
\label{fig:SC4}
\end{figure}

>>>>

\includegraphics[scale=0.8]{figures/fig1/fig1.pdf}

@@

\begin{figure}
\vspace{-1cm}
\begin{center}
\includegraphics[scale=.8]{figures/fig2/fig2.pdf}
\end{center}
\vspace{-1cm}
\caption{\small Emergent property inference of a stochastic stabilized supralinear network.
\textbf{A}.  Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).   
Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_b$.
Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
\textbf{B}. Responses of the deterministic model ($\bm{\epsilon}=\bm{0}$) to varying contrasts.  The response at 50\% contrast (dots) is the focus of our analysis.
\textbf{C}. Paradoxical response of the stochastic model to a small increase in input to the P-population.  
\textbf{D}. EPI posterior of noise parameters $\mathbf{z}$ conditioned on realistic E-population Fano factors.
The posterior predictive distribution is show on the bottom-left.
and the mode of the distribution is starred.
\textbf{E}. (Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of the posterior.
Each gray dot is a choice of $\sigma_P$, for which a constrained mode $z^*(\sigma_P, P)$ is chosen.
The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
(Bottom) Such sensitive dimensions of the Hessian (dots) are significantly more sensitive than randomly chosen dimensions (box and whiskers).
\textbf{F}. The Fano factor of the E-population is strongly correlated with each other neuron-type population.
\textbf{G}. Mean and standard deviation (across EPI posterior) of Fano factor of each neuron-type population at each level of contrast.
}
 \label{fig:V1}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
\end{center}
\caption{\small 
\textbf{A}. Rapid task switching behavioral paradigm (see text). 
\textbf{B}. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
\textbf{C}. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice of simulations (D).
\textbf{D}. Simulations of an SC network from the EPI distribution with 75\% accuracy in each task.  Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (optogenetic strength $\gamma=0.7$).  Shading indicates standard deviation across trials.
\textbf{E}. Difference in performance of each task during inactivation.  Inactivation level $\gamma$ scales from no inactivation (0) to full inactivation (1).  We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
\textbf{F}. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
}
\label{fig:SC}
\end{figure}

@@

Here, we are interested in the scalability of EPI in number of parameters $|\mathbf{z}|$.  We consider rank-2 RNN with N neurons of connectivity
\begin{equation}
 W = UV^\top + g\chi
 \end{equation}
 and dynamics
 \begin{equation}
 \tau \dot{\mathbf{x}} = -\mathbf{x} + W\mathbf{x}
 \end{equation}
where $U = \begin{bmatrix} \mathbf{u}_1 & \mathbf{u}_2 \end{bmatrix}$, $V = \begin{bmatrix} \mathbf{v}_1 & \mathbf{v}_2 \end{bmatrix}$, $\mathbf{u}_1, \mathbf{u}_2, \mathbf{v}_1, \mathbf{v}_2 \in \left[-1, 1 \right]^N$, and $g=0.01$.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=.8]{figures/fig4/fig4.pdf}
\end{center}
\caption{\small 
\textbf{A}. Recurrent neural network.
\textbf{B}. EPI scales with $z$ to high dimensions. 
Convergence definitions: 
EPI (blue) - satisfies all moment constraints, 
SNPE (orange)- produces at least 2/$n_{\text{train}}$ parameter samples are in the bounds of emergent property (mean +- 0.5), 
and SMC-ABC (red) - 100 particles with $\epsilon < 0.5$ are produced.
\textbf{C}. Posterior predictive distributions of EPI (blue), SNPE (orange), and SMC-ABC (red). 
Gray star indicates emergent property mean, and gray dashed lines indicate two standard deviations corresponding to the variance constraint.
For $N <= 6$ where SMC-ABC converges, samples are not diverse (path degeneracies).  
For $N >= 25$, SNPE does not produce a posterior approximation yielding parameters with simulations near $x_0$. 
\textbf{D}. Simulations of network parameters resulting from each method ($\tau=100ms$).  Each trace corresponds to simulation of one $z$.  
\textbf{E}. Ratio of obtained samples producing stable amplification.
}
\label{fig:LRRNN}
\end{figure}

@@

However, this computation has cubic complexity in dimensionality for fully connected layers.  
By restricting our layers to normalizing flows \cite{rezende2015variational, papamakarios2019normalizing} -- bijective functions with fast log determinant Jacobian computations, which confer a fast calculation of the sample log probability.
Fast log probability calculation confers efficient optimization of the maximum entropy objective (see Section \ref{methods_AL_opt}).
%Most of our analyses use either a planar flow \cite{rezende2015variational} or real NVP \cite{dinh2017density}, which have proven effective in our architecture searches.  
%Planar flow architectures are specified by the number of planar bijection layers used, while real NVP architectures are specified by the number of masks, neural network layers per mask, units per layer, and batch normalization momentum parameter.
We use the Real NVP \cite{dinh2017density} normalizing flow class, because its coupling architecture confers both fast sampling (forward) and fast log probability evaluation (backward).
Fast probability evaluation in turn facilitates fast gradient and Hessian evaluation of log probability throughout parameter space.
Glow permutations were used in between coupling stages \cite{kingma2018glow}.
This is in contrast to autoregressive architectures \cite{papamakarios2017masked, kingma2016improved}, in which only forward or backward passes are efficient.

In this work, normalizing flows are used as flexible posterior approximations $q_{\bm{\theta}}(\mathbf{z})$ having weights and biases $\bm{\theta}$. We specify the architecture used in each application by the number of Real-NVP affine coupling stages, and the number of neural network layers and units per layer of the conditioning functions.

\subsubsection{Augmented Lagrangian optimization}\label{methods_AL_opt}
To optimize $q_{\bm{\theta}}(\mathbf{z})$ in Equation \ref{eq:opt}, the constrained maximum entropy optimization is executed using the augmented Lagrangian method.  The following objective is minimized:
\begin{equation} \label{eq:AL}
L(\bm{\theta}; \bm{\eta}_{\text{opt}}, c) = -H(q_{\bm{\theta}}) + \bm{\eta}_{\text{opt}}^\top R(\bm{\theta}) + \frac{c}{2}||R(\bm{\theta})||^2
\end{equation}
where average constraint violations $R(\bm{\theta}) = \mathbb{E}_{\mathbf{z} \sim q_{\bm{\theta}}}\left[ \mathbb{E}_{\mathbf{x}\sim p(\mathbf{x} \mid \mathbf{z})}\left[T(\mathbf{x}; \mathbf{z}) - \bm{\mu}_{\text{opt}} \right] \right]$, $\bm{\eta}_{\text{opt}} \in \mathbb{R}^m$ are the Lagrange multipliers where $m = |\bm{\mu}_{\text{opt}}| = |T(\mathbf{x}; \mathbf{z})| = 2|f(\mathbf{x}; \mathbf{z})|$,  and $c$ is the penalty coefficient. 
The sufficient statistics $T(\mathbf{x}; \mathbf{z})$ and mean parameter $\bm{\mu}_{\text{opt}}$ are determined by the means $\bm{\mu}$ and variances $\bm{\sigma}^2$ of emergent property statistics $f(\mathbf{x}; \mathbf{z})$ defined in Equation \ref{eq:opt}.
Specifically, $T(\mathbf{x}; \mathbf{z})$ is a concatenation of the first and second moments, $\bm{\mu}_{\text{opt}}$ is a concatenation of $\bm{\mu}$ and $\bm{\sigma}^2$ (see section \ref{methods_ME_EF}), and the Lagrange multipliers are closely related to the natural parameters $\bm{\eta}$ of exponential families (see Section \ref{methods_VI}).
Weights and biases $\bm{\theta}$ of the deep probability distribution are optimized according to Equation \ref{eq:AL} using the Adam optimizer with learning rate $10^{-3}$ \cite{kingma2014adam}.

@@

The lagrangian parameters $\bm{\eta}_{\text{opt}}$ are initialized to zero and adapted following each augmented Lagrangian epoch, which is a period of optimization with fixed ($\bm{\eta}_{\text{opt}}$, $c$) for a given number of stochastic optimization iterations. 
A low value of $c$ is used initially, and conditionally increased after each epoch based on constraint error reduction.
The penalty coefficient is updated based on the result of a hypothesis test regarding the reduction in constraint violation.  
The p-value of $\mathbb{E}[||R(\bm{\theta}_{k+1})||] > \gamma \mathbb{E} \left[||R(\bm{\theta}_{k})|| \right]$ is computed, and $c_{k+1}$ is updated  to $\beta c_k$ with probability $1-p$.  
The other update rule is $\bm{\eta}_{\text{opt},k+1} = \bm{\eta}_{\text{opt},k} + c_k \frac{1}{n} \sum_{i=1}^n (T(\mathbf{x}^{(i)}) - \bm{\mu}_{\text{opt}})$ given a batch size $n$.
Throughout the study, $\gamma = 0.25$, while $\beta$ was chosen to be either 2 or 4.  The batch size of EPI also varied according to application.

@@

To assess whether the EPI distribution $q_{\bm{\theta}}(\mathbf{z})$ produces the emergent property, we assess whether each individual constraint on the means and variances of $f(\mathbf{x}; \mathbf{z})$ is satisfied.
We consider the EPI to have converged when a null hypothesis test of constraint violations $R(\bm{\theta})_i$ being zero is accepted for all constraints $i \in \{1, ..., m\}$ at a significance threshold $\alpha=0.05$.  
This significance threshold is adjusted through Bonferroni correction according to the number of constraints $m$.  
The p-values for each constraint are calculated according to a two-tailed nonparametric test, where 200 estimations of the sample mean $R(\bm{\theta})^i$ are made using $N_{\text{test}}$ samples of $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ at the end of the augmented Lagrangian epoch.

@@

\subsubsection{Example: 2D LDS}\label{methods_2DLDS}
To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
\begin{equation} 
\tau \frac{d\mathbf{x}}{dt} = A\mathbf{x}
\end{equation}
with
\begin{equation}
A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
\end{equation}
To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(\mathbf{x})$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system. 
 $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
\begin{equation}
 \mathbb{E}\left[T(\mathbf{x}) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}.
 \end{equation} 
\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{figures/figLDS1/figLDS1.pdf}
\end{center}
\begin{flushleft}
\caption{\small (LDS1): A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
B. The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.  Dashed lines indicate the parameter axes. 
C. Entropy throughout the optimization.  
At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.  
D. Emergent property moments throughout optimization.  
At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
\end{flushleft}
\label{fig:LDS1}
\end{figure}

@@

For EPI in Fig \ref{fig:STG}E, we used a real NVP architecture with three Real NVP coupling layers and two-layer neural networks of 25 units per layer.
The initial distribution was a standard isotropic gaussian $z_0 \sim \mathcal{N}(\mathbf{0}, I)$ mapped to a support of $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}] \in [4,8] \times [0.01,4]$. 
We did not include $g_{\text{synA}} < 0.01$, since conductances that low make the circuit simulations numerically unstable.
We used an augmented Lagrangian coefficient of $c_0 = 10^{5}$, a batch size $n=400$, $\beta = 2$, $N_{\text{test}} = 100$, and initialized $q_{\bm{\theta}}(\mathbf{z})$ to produce a gaussian approximation to samples returned from an initial ABC search.
This initialization had much greater entropy and a different emergent property than the returned EPI posterior.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{figures/figSC1/figSC1.pdf}
\end{center}
\caption{\small (SC1): Connectivity parameters of EPI distributions versus task accuracies.  $\beta$ is slope coefficient of linear regression, $r$ is correlation, and $p$ is the two-tailed p value.
}
\label{fig:SC1}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=.7]{figures/figSC2/figSC2.pdf}
\end{center}
\caption{\small (SC2):  
A. Invariant eigenvectors of connectivity matrix $W$.
B. Eigenvalues of connectivities of EPI distribution versus task accuracies.}
\label{fig:SC2}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=1.]{figures/figSC3/figSC3.pdf}
\end{center}
\caption{\small (SC3): 
A. Connectivity eigenvalues of EPI parameter distribution colored by Pro task accuracy. 
B. Same for Anti task.
}
\label{fig:SC3}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{figures/figSC4/figSC4.pdf}
\end{center}
\caption{\small (SC4):  Scatters of the effect of delay period inactivation in each task with task accuracy.  Plots are shown at an opto strength of 0.8.
}
\label{fig:SC4}
\end{figure}
